---
title: "EORTC HNSCC ICI Honore et al_02032025"
output: html_notebook
---
library(swimplot)
library(grid)
library(gtable)
library(readr) 
library(mosaic)
library(dplyr) 
library(survival) 
library(survminer) 
library(ggplot2)
library(scales)
library(coxphf)
library(ggthemes)
library(tidyverse)
library(gtsummary)
library(flextable)
library(parameters)
library(car)
library(ComplexHeatmap)
library(tidyverse)
library(readxl)
library(janitor)
library(DT)
library(pROC)
library(rms)

#ctDNA Detection Rates by Window and Stages
```{r}
#ctDNA at Baseline
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data$ctDNA.Base <- factor(circ_data$ctDNA.Base, levels=c("NEGATIVE","POSITIVE"))
circ_data <- subset(circ_data, ctDNA.Base %in% c("NEGATIVE", "POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("O","I","II","III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.Base == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.Base, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.Base == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)

#ctDNA post-treatment
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.postTx!="",]
circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels=c("NEGATIVE","POSITIVE"))
circ_data$Stage <- factor(circ_data$Stage, levels=c("O","I","II","III","IV"))
positive_counts_by_stage <- aggregate(circ_data$ctDNA.postTx == "POSITIVE", by=list(circ_data$Stage), FUN=sum)
total_counts_by_stage <- aggregate(circ_data$ctDNA.postTx, by=list(circ_data$Stage), FUN=length)
combined_data <- data.frame(
  Stage = total_counts_by_stage$Group.1,
  Total_Count = total_counts_by_stage$x,
  Positive_Count = positive_counts_by_stage$x,
  Rate = (positive_counts_by_stage$x / total_counts_by_stage$x) * 100  # Convert to percentage
)
combined_data$Rate <- sprintf("%.2f%%", combined_data$Rate)
overall_total_count <- nrow(circ_data)
overall_positive_count <- nrow(circ_data[circ_data$ctDNA.postTx == "POSITIVE",])
overall_positivity_rate <- (overall_positive_count / overall_total_count) * 100  # Convert to percentage
overall_row <- data.frame(
  Stage = "Overall",
  Total_Count = overall_total_count,
  Positive_Count = overall_positive_count,
  Rate = sprintf("%.2f%%", overall_positivity_rate)
)
combined_data <- rbind(combined_data, overall_row)
print(combined_data)
```



#Overview plot
```{r}
setwd("~/Downloads") 
clinstage <- read.csv("EORTC ICI_OP.csv")
clinstage_df <- as.data.frame(clinstage)

# Creating the basic swimmer plot
oplot <- swimmer_plot(df=clinstage_df,
                      id='PatientName',
                      end='fu.diff.months',
                      fill='gray',
                      width=.01)

# Adding themes and scales
oplot <- oplot + theme(panel.border = element_blank())
oplot <- oplot + scale_y_continuous(breaks = seq(0, 48, by = 3))
oplot <- oplot + labs(x ="Patients", y="Months from Immunotherapy Start")

# Adding swimmer points
oplot_ev1 <- oplot + swimmer_points(df_points=clinstage_df,
                                    id='PatientName',
                                    time='date.diff.months',
                                    name_shape ='Event_type',
                                    name_col = 'Event',
                                    size=3.5,fill='black')
# Optionally uncomment and use col='darkgreen' if needed

# Adding shape manual scale
oplot_ev1.1 <- oplot_ev1 + ggplot2::scale_shape_manual(name="Event_type",
                                                       values=c(1,16,6,4),
                                                       breaks=c('ctDNA_neg','ctDNA_pos','Imaging','Death'))

# Display the plot
oplot_ev1.1
                                    oplot_ev2 <- oplot_ev1.1 + swimmer_lines(df_lines=clinstage_df,
                                                                             id='PatientName',
                                                                             start='Tx_start.months',
                                                                             end='Tx_end.months',
                                                                             name_col='Tx_type',
                                                                             size=3.5,
                                                                             name_alpha = 1.0)
                                    oplot_ev2 <- oplot_ev2 + guides(linetype = guide_legend(override.aes = list(size = 5, color = "black")))
                                    oplot_ev2
oplot_ev2.2 <- oplot_ev2 + ggplot2::scale_color_manual(name="Event",values=c( "orange", "black", "black", "lightblue", "green", "red"))
oplot_ev2.2
```

#Overview plot - stratified by BOR
```{r}
setwd("~/Downloads") 
clinstage <- read.csv("EORTC ICI_OP.csv")
clinstage_df <- as.data.frame(clinstage)

# Creating the basic swimmer plot
oplot_stratify <-swimmer_plot(df=clinstage_df,
                              id='PatientName',
                              end='fu.diff.months',
                              col="gray",
                              alpha=0.75,
                              width=.01,
                              base_size = 14,
                              stratify= c('RECIST'))
oplot_stratify <- oplot_stratify + theme(panel.border = element_blank())
oplot_stratify <- oplot_stratify + scale_y_continuous(breaks = seq(0, 42, by = 3))
oplot_stratify <- oplot_stratify + labs(x ="Patients" , y="Months from Immunotherapy Start")

# Adding swimmer points
oplot_ev1 <- oplot_stratify + swimmer_points(df_points=clinstage_df,
                                    id='PatientName',
                                    time='date.diff.months',
                                    name_shape ='Event_type',
                                    name_col = 'Event',
                                    size=3.5,fill='black')
# Optionally uncomment and use col='darkgreen' if needed

# Adding shape manual scale
oplot_ev1.1 <- oplot_ev1 + ggplot2::scale_shape_manual(name="Event_type",
                                                       values=c(1,16,6,4),
                                                       breaks=c('ctDNA_neg','ctDNA_pos','Imaging','Death'))

# Display the plot
oplot_ev1.1
                                    oplot_ev2 <- oplot_ev1.1 + swimmer_lines(df_lines=clinstage_df,
                                                                             id='PatientName',
                                                                             start='Tx_start.months',
                                                                             end='Tx_end.months',
                                                                             name_col='Tx_type',
                                                                             size=3.5,
                                                                             name_alpha = 1.0)
                                    oplot_ev2 <- oplot_ev2 + guides(linetype = guide_legend(override.aes = list(size = 5, color = "black")))
                                    oplot_ev2
oplot_ev2.2 <- oplot_ev2 + ggplot2::scale_color_manual(name="Event",values=c( "orange", "black", "black", "lightblue", "green", "red"))
oplot_ev2.2
```

#Association of Baseline ctDNA MTM levels with clinicopathological factors
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_datadf <- as.data.frame(circ_data)

tally(~cT.Status, data=circ_data, margins = TRUE)
circ_data$cT.Status <- factor(circ_data$cT.Status, levels = c("cT0-T2","cT3-T4"), labels = c("cT0-T2 (n=11)","cT3-T4 (n=17)"))
boxplot(ctDNA.Base.MTM~cT.Status, data=circ_data, main="ctDNA pre-treatment MTM - cT status", xlab="cT status", ylab="MTM/mL", col="white",border="black")
m3<-wilcox.test(ctDNA.Base.MTM ~ cT.Status, data=circ_data, na.rm=TRUE, exact=FALSE, conf.int=TRUE)
print(m3)

tally(~cN.Status, data=circ_data, margins = TRUE)
circ_data$cN.Status <- factor(circ_data$cN.Status, levels = c("cN0","cN1-N3"), labels = c("cN0 (n=12)","cN1-N3 (n=15)"))
boxplot(ctDNA.Base.MTM~cN.Status, data=circ_data, main="ctDNA pre-treatment MTM - cN status", xlab="cN status", ylab="MTM/mL", col="white",border="black")
m4<-wilcox.test(ctDNA.Base.MTM ~ cN.Status, data=circ_data, na.rm=TRUE, exact=FALSE, conf.int=TRUE)
print(m4)
```

#Median MTM/mL levels for ctDNA positive pts pre-treatment
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]

median_ctDNA <- median(circ_data$ctDNA.Base.MTM, na.rm = TRUE)
range_ctDNA <- range(circ_data$ctDNA.Base.MTM, na.rm = TRUE)
cat("Median MTM/mL post-treatment:", median_ctDNA, "\n")
cat("Range MTM/mL post-treatment:", range_ctDNA[1], "-", range_ctDNA[2], "\n")
```

#Median MTM/mL levels for ctDNA positive pts post-treatment
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.postTx!="",]
circ_data <- circ_data[circ_data$ctDNA.postTx=="POSITIVE",]

median_ctDNA <- median(circ_data$ctDNA.postTx.MTM, na.rm = TRUE)
range_ctDNA <- range(circ_data$ctDNA.postTx.MTM, na.rm = TRUE)
cat("Median MTM/mL post-treatment:", median_ctDNA, "\n")
cat("Range MTM/mL post-treatment:", range_ctDNA[1], "-", range_ctDNA[2], "\n")
```

#Median time from end treatment to progression for ctDNA negative pts post-treatment
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.postTx!="",]
circ_data <- circ_data[circ_data$ctDNA.postTx=="NEGATIVE",]
circ_data <- circ_data[circ_data$PFS.Event=="TRUE",]

median_PFS <- median(circ_data$PFS.months, na.rm = TRUE)
range_PFS <- range(circ_data$PFS.months, na.rm = TRUE)
cat("Median PFS:", median_PFS, "\n")
cat("Range PFS:", range_PFS[1], "-", range_PFS[2], "\n")
```

#PFS by ctDNA status post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.postTx!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)~ctDNA.postTx, data = circ_data)
event_summary <- circ_data %>%
  group_by(ctDNA.postTx) %>%
  summarise(
    Total = n(),
    Events = sum(PFS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.postTx, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="PFS - ctDNA Post/Under treatment", ylab= "Progression-Free Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
summary(KM_curve, times= c(0, 12, 24, 36))
circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ctDNA.postTx + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels = c("NEGATIVE", "POSITIVE"), labels = c("Negative", "Positive"))
circ_data$PFS.Event <- factor(circ_data$PFS.Event, levels = c("FALSE", "TRUE"), labels = c("No Progression", "Progression"))
contingency_table <- table(circ_data$ctDNA.postTx, circ_data$PFS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA status post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Progression",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("No Progression" = "blue", "Progression" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#OS by ctDNA status post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.postTx!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$OS.months, event = circ_data$OS.Event)~ctDNA.postTx, data = circ_data)
event_summary <- circ_data %>%
  group_by(ctDNA.postTx) %>%
  summarise(
    Total = n(),
    Events = sum(OS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.postTx, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="OS - ctDNA Post/Under treatment", ylab= "Overall Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("ctDNA Negative", "ctDNA Positive"), legend.title="")
summary(KM_curve, times= c(12, 24, 36))
circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ctDNA.postTx + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels = c("NEGATIVE", "POSITIVE"), labels = c("Negative", "Positive"))
circ_data$OS.Event <- factor(circ_data$OS.Event, levels = c("FALSE", "TRUE"), labels = c("Alive", "Deceased"))
contingency_table <- table(circ_data$ctDNA.postTx, circ_data$OS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA status post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Living Status",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Alive" = "blue", "Deceased" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#Association of ctDNA status post/during-ICI with BOR
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.postTx <- factor(circ_data$ctDNA.postTx, levels = c("NEGATIVE", "POSITIVE"), labels = c("Negative", "Positive"))
circ_data$RESIST <- factor(circ_data$RESIST, levels = c("CR", "PR", "SD", "PD"))
contingency_table <- table(circ_data$ctDNA.postTx, circ_data$RESIST)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA status post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "BOR",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("CR" = "blue", "PR" = "lightblue", "SD" = "orange", "PD" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#PFS by ctDNA kinetics post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)~ΔctDNA, data = circ_data)
event_summary <- circ_data %>%
  group_by(ΔctDNA) %>%
  summarise(
    Total = n(),
    Events = sum(PFS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
KM_curve <- survfit(surv_object ~ ΔctDNA, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="PFS - ctDNA Kinetics Post/Under treatment", ylab= "Progression-Free Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("Decreased ΔctDNA", "Increased ΔctDNA"), legend.title="")
summary(KM_curve, times= c(0, 12, 24, 36))
circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ΔctDNA + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels = c("NEGATIVE", "POSITIVE"), labels = c("Decreased", "Increased"))
circ_data$PFS.Event <- factor(circ_data$PFS.Event, levels = c("FALSE", "TRUE"), labels = c("No Progression", "Progression"))
contingency_table <- table(circ_data$ΔctDNA, circ_data$PFS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA kinetics post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Progression",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("No Progression" = "blue", "Progression" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#OS by ctDNA kinetics post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
circ_datadf <- as.data.frame(circ_data)

survfit(Surv(time = circ_data$OS.months, event = circ_data$OS.Event)~ΔctDNA, data = circ_data)
event_summary <- circ_data %>%
  group_by(ΔctDNA) %>%
  summarise(
    Total = n(),
    Events = sum(OS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
KM_curve <- survfit(surv_object ~ ΔctDNA, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="OS - ctDNA Kinetics Post/Under treatment", ylab= "Overall Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("Decreased ΔctDNA", "Increased ΔctDNA"), legend.title="")
summary(KM_curve, times= c(0, 12, 24, 36))
circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ΔctDNA + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels = c("NEGATIVE", "POSITIVE"), labels = c("Decreased", "Increased"))
circ_data$OS.Event <- factor(circ_data$OS.Event, levels = c("FALSE", "TRUE"), labels = c("Alive", "Deceased"))
contingency_table <- table(circ_data$ΔctDNA, circ_data$OS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA kinetics post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Vital Status",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Alive" = "blue", "Deceased" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#Association of ctDNA kinetics post/during-ICI with BOR
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels = c("NEGATIVE", "POSITIVE"), labels = c("Decreased", "Increased"))
circ_data$RESIST <- factor(circ_data$RESIST, levels = c("CR", "PR", "SD", "PD"))
contingency_table <- table(circ_data$ΔctDNA, circ_data$RESIST)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA kinetics post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "BOR",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("CR" = "blue", "PR" = "lightblue", "SD" = "orange", "PD" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#PFS by ctDNA clearance post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))

survfit(Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)~ctDNA.Dynamics, data = circ_data)
event_summary <- circ_data %>%
  group_by(ctDNA.Dynamics) %>%
  summarise(
    Total = n(),
    Events = sum(PFS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.Dynamics, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="PFS - ctDNA clearance Post/Under treatment", ylab= "Progression-Free Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("Clearance", "No Clearance"), legend.title="")
summary(KM_curve, times= c(0, 12, 24, 36))
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$PFS.Event <- factor(circ_data$PFS.Event, levels = c("FALSE", "TRUE"), labels = c("No Progression", "Progression"))
contingency_table <- table(circ_data$ctDNA.Dynamics, circ_data$PFS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA clearance post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Progression",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("No Progression" = "blue", "Progression" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#OS by ctDNA clearance post/during-ICI
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))

survfit(Surv(time = circ_data$OS.months, event = circ_data$OS.Event)~ctDNA.Dynamics, data = circ_data)
event_summary <- circ_data %>%
  group_by(ctDNA.Dynamics) %>%
  summarise(
    Total = n(),
    Events = sum(OS.Event),
    Fraction = Events / n(),
    Percentage = (Events / n()) * 100
  )
print(event_summary)
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
KM_curve <- survfit(surv_object ~ ctDNA.Dynamics, data = circ_data,conf.int=0.95,conf.type="log-log") 
ggsurvplot(KM_curve, data = circ_data, pval = FALSE, conf.int = FALSE, risk.table = TRUE, break.time.by=6, palette=c("blue","red"), title="OS - ctDNA clearance Post/Under treatment", ylab= "Overall Survival", xlab="Months from Start of Immunotherapy", legend.labs=c("Clearance", "No Clearance"), legend.title="")
summary(KM_curve, times= c(0, 12, 24, 36))
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics + Inclusion.status + p16.status + ECOG.v2, data=circ_data) 
ggforest(cox_fit,data = circ_data) 
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

circ_data$OS.Event <- factor(circ_data$OS.Event, levels = c("FALSE", "TRUE"), labels = c("Alive", "Deceased"))
contingency_table <- table(circ_data$ctDNA.Dynamics, circ_data$OS.Event)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA clearance post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "Vital Status",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("Alive" = "blue", "Deceased" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#Association of ctDNA clearance post/during-ICI with BOR
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))

circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$RESIST <- factor(circ_data$RESIST, levels = c("CR", "PR", "SD", "PD"))
contingency_table <- table(circ_data$ctDNA.Dynamics, circ_data$RESIST)
chi_square_test <- chisq.test(contingency_table)
print(chi_square_test)
fisher_exact_test <- fisher.test(contingency_table)
print(fisher_exact_test)
print(contingency_table)
table_df <- as.data.frame(contingency_table)
table_df$Total <- ave(table_df$Freq, table_df$Var1, FUN = sum)
table_df$Percentage <- table_df$Freq / table_df$Total
table_df$MiddlePercentage <- table_df$Percentage / 2
ggplot(table_df, aes(x = Var1, y = Percentage, fill = Var2)) +
  geom_bar(stat = "identity") +
  geom_text(aes(y = MiddlePercentage, label = Freq), position = "stack", color = "black", vjust = 1.5, size = 7) +
  theme_minimal() +
  labs(title = "ctDNA clearance post/under-treatment", 
       x = "ctDNA", 
       y = "Patients (%)", 
       fill = "BOR",
       caption = paste("Fisher's exact test p-value: ", format.pval(fisher_exact_test$p.value))) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = c("CR" = "blue", "PR" = "lightblue", "SD" = "orange", "PD" = "red")) + # define custom colors
  theme(axis.text.x = element_text(angle = 0, hjust = 1.5, size = 14), # increase x-axis text size
        axis.text.y = element_text(size = 14, color = "black"), # increase y-axis text size
        axis.title.x = element_text(size = 14, color = "black"), # increase x-axis label size
        axis.title.y = element_text(size = 14, color = "black"), # increase y-axis label size
        legend.text = element_text(size = 12, color = "black"))  # increase Progression label size
```

#Multivariate cox regression for PFS - ctDNA clearance
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))

circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$CPS.Scorev2 <- factor(circ_data$CPS.Scorev2, levels = c("≥1", "<1"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
surv_object <- Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event) 
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics + ECOG.v2 + Inclusion.status + p16.status + Immunotherapy, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for PFS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```


#Multivariate cox regression for PFS - ctDNA kinetics
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$CPS.Scorev2 <- factor(circ_data$CPS.Scorev2, levels = c("≥1", "<1"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
surv_object <- Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event) 
cox_fit <- coxph(surv_object ~ ΔctDNA + ECOG.v2 + Inclusion.status + p16.status + Immunotherapy, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for PFS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```


#Univariate PFS cox regression for variables included in MVA
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance")) #univariate for ctDNA clearance post-treatment
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
cox_fit <- coxph(surv_object ~ ΔctDNA, data=circ_data) #univariate for ΔctDNA
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ECOG.v2, data=circ_data) #univariate for ECOG
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic")) #univariate for Stage/disease status
cox_fit <- coxph(surv_object ~ Inclusion.status, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive")) #univariate for p16 status
cox_fit <- coxph(surv_object ~ p16.status, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$PFS.months, event = circ_data$PFS.Event)
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
cox_fit <- coxph(surv_object ~ Immunotherapy, data=circ_data) #univariate for Immunotherapy regimen
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```

#Multivariate cox regression for OS - ctDNA clearance
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))

circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$CPS.Scorev2 <- factor(circ_data$CPS.Scorev2, levels = c("≥1", "<1"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
surv_object <- Surv(time = circ_data$OS.months, event = circ_data$OS.Event) 
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics + ECOG.v2 + Inclusion.status + p16.status + Immunotherapy, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for OS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```

#Multivariate cox regression for OS - ctDNA kinetics
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic"))
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive"))
circ_data$CPS.Scorev2 <- factor(circ_data$CPS.Scorev2, levels = c("≥1", "<1"))
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
surv_object <- Surv(time = circ_data$OS.months, event = circ_data$OS.Event) 
cox_fit <- coxph(surv_object ~ ΔctDNA + ECOG.v2 + Inclusion.status + p16.status + Immunotherapy, data=circ_data) 
ggforest(cox_fit, data = circ_data, main = "Multivariate Regression Model for OS", refLabel = "Reference Group")
test.ph <- cox.zph(cox_fit)
```

#Univariate OS cox regression for variables included in MVA
```{r}
rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance")) #univariate for ctDNA clearance post-treatment
cox_fit <- coxph(surv_object ~ ctDNA.Dynamics, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ΔctDNA!="",]
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$ΔctDNA <- factor(circ_data$ΔctDNA, levels=c("NEGATIVE","POSITIVE"))
cox_fit <- coxph(surv_object ~ ΔctDNA, data=circ_data) #univariate for ΔctDNA
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$ECOG.v2 <- factor(circ_data$ECOG.v2, levels = c("0", "1/2"))
cox_fit <- coxph(surv_object ~ ECOG.v2, data=circ_data) #univariate for ECOG
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$Inclusion.status <- factor(circ_data$Inclusion.status, levels = c("Loco-regional", "Metastatic")) #univariate for Stage/disease status
cox_fit <- coxph(surv_object ~ Inclusion.status, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$p16.status <- factor(circ_data$p16.status, levels = c("Negative", "Positive")) #univariate for p16 status
cox_fit <- coxph(surv_object ~ p16.status, data=circ_data)
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)

rm(list=ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
surv_object <-Surv(time = circ_data$OS.months, event = circ_data$OS.Event)
circ_data$Immunotherapy <- factor(circ_data$Immunotherapy, levels = c("Pembrolizumab", "Nivolumab"))
cox_fit <- coxph(surv_object ~ Immunotherapy, data=circ_data) #univariate for Immunotherapy regimen
summary(cox_fit)
cox_fit_summary <- summary(cox_fit)

#Extract values for HR, 95% CI, and p-value
HR <- cox_fit_summary$coefficients[2]
lower_CI <- cox_fit_summary$conf.int[3]
upper_CI <- cox_fit_summary$conf.int[4]
p_value <- cox_fit_summary$coefficients[5]
label_text <- paste0("HR = ", round(HR, 2), " (", round(lower_CI, 2), "-", round(upper_CI, 2), "); p = ", round(p_value, 3))
print(label_text)
```

#Waterfall with % change for MTM/mL
```{r}
rm(list = ls())
setwd("~/Downloads") 
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available == "TRUE", ]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data$ctDNA.change <- as.numeric(circ_data$ctDNA.change)
circ_data$SubjectID <- seq_len(nrow(circ_data))
circ_data <- circ_data %>% arrange(desc(ctDNA.change))
circ_data$SubjectID <- factor(circ_data$SubjectID, levels = circ_data$SubjectID)

ggplot(circ_data, aes(x = SubjectID, y = ctDNA.change)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = 0, color = "black") +
  scale_y_continuous(
    breaks = seq(-1000, 1000, by = 200),
    limits = c(-1000, 1000)
  ) +
  labs(
    x = "Individual (Sorted)",
    y = "ctDNA % Change",
    title = "ctDNA Change During ICI per Individual"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, size = 6),
    plot.title = element_text(hjust = 0.5)
  )
```


#Heatmap for the clinical factors and top 10 most common mutations
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data$ctDNA.change <- as.numeric(circ_data$ctDNA.change)
circ_data <- circ_data %>% arrange(desc(ctDNA.change))
circ_datadf <- as.data.frame(circ_data)

ha <- HeatmapAnnotation(
  Inclusion.status = circ_data$Inclusion.status,
  Immunotherapy = circ_data$Immunotherapy,
  ECOG.v2 = circ_data$ECOG.v2,
  p16.status = circ_data$p16.status,
  TP53 = circ_data$TP53,
  CDKN2A = circ_data$CDKN2A,
  FAT1 = circ_data$FAT1,
  APOB = circ_data$APOB,
  LAMA1 = circ_data$LAMA1,
  NOTCH1 = circ_data$NOTCH1,
  PCDHB13 = circ_data$PCDHB13,
  DCHS1 = circ_data$DCHS1,
  FSIP2 = circ_data$FSIP2,
  LRP1B = circ_data$LRP1B,
  ctDNA.Base = circ_data$ctDNA.Base,
  ctDNA.postTx = circ_data$ctDNA.postTx,
  RESIST = circ_data$RESIST,
  PFS.Event = circ_data$PFS.Event,
  OS.Event = circ_data$OS.Event,
  
  col = list(Inclusion.status = c("Loco-regional" = "seagreen1", "Metastatic" = "khaki"),
             Immunotherapy = c("Pembrolizumab" = "purple", "Nivolumab" ="lightblue"),           
             ECOG.v2 = c("0" = "goldenrod" , "1/2" = "blue4"),
             p16.status = c("Negative" = "yellow", "Positive" ="brown", "Unknown" ="grey"),
             TP53 = c("MUT" = "black", "WT" ="grey"),
             CDKN2A = c("MUT" = "black", "WT" ="grey"),
             FAT1 = c("MUT" = "black", "WT" ="grey"),
             APOB = c("MUT" = "black", "WT" ="grey"),
             LAMA1 = c("MUT" = "black", "WT" ="grey"),
             NOTCH1 = c("MUT" = "black", "WT" ="grey"),
             PCDHB13 = c("MUT" = "black", "WT" ="grey"),
             DCHS1 = c("MUT" = "black", "WT" ="grey"),
             FSIP2 = c("MUT" = "black", "WT" ="grey"),
             LRP1B = c("MUT" = "black", "WT" ="grey"),
             ctDNA.Base = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
             ctDNA.postTx = c("POSITIVE" = "red3", "NEGATIVE" ="blue"),
             RESIST = c("CR" = "lightblue", "PR" ="blue", "SD" = "orange", "PD" ="red"),
             PFS.Event = c("TRUE" = "red3", "FALSE" ="blue"),
             OS.Event = c("TRUE" = "black", "FALSE" ="grey")
  )
)
ht <- Heatmap(matrix(nrow = 0, ncol = length(circ_data$Inclusion.status)),show_row_names = FALSE,cluster_rows = F,cluster_columns = FALSE, top_annotation = ha)
pdf("heatmap.pdf",width = 7, height = 7)
draw(ht, annotation_legend_side = "bottom")
dev.off()
```

#Frequency of top 10 mutations in ctDNA clearance vs non-clearance
```{r}
rm(list=ls())
setwd("~/Downloads")
circ_data <- read.csv("EORTC HNSCC ICI Clinical Data.csv")
circ_data <- circ_data[circ_data$ctDNA.available=="TRUE",]
circ_data <- circ_data[circ_data$ctDNA.Base!="",]
circ_data <- circ_data[circ_data$ctDNA.Base=="POSITIVE",]
circ_datadf <- as.data.frame(circ_data)

circ_data$ctDNA.Dynamics <- NA
circ_data <- circ_data %>%
  mutate(ctDNA.Dynamics = case_when(
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "NEGATIVE" ~ 1,
    ctDNA.Base == "POSITIVE" & ctDNA.postTx == "POSITIVE" ~ 2
  ))
circ_data$ctDNA.Dynamics <- factor(circ_data$ctDNA.Dynamics, levels=c("1","2"), labels = c("Clearance", "No Clearance"))
circ_data$TP53 <- factor(circ_data$TP53, levels=c("WT","MUT"))
circ_data$CDKN2A <- factor(circ_data$CDKN2A, levels=c("WT","MUT"))
circ_data$FAT1 <- factor(circ_data$FAT1, levels=c("WT","MUT"))
circ_data$APOB <- factor(circ_data$APOB, levels=c("WT","MUT"))
circ_data$LAMA1 <- factor(circ_data$LAMA1, levels=c("WT","MUT"))
circ_data$NOTCH1 <- factor(circ_data$NOTCH1, levels=c("WT","MUT"))
circ_data$PCDHB13 <- factor(circ_data$CUBN, levels=c("WT","MUT"))
circ_data$DCHS1 <- factor(circ_data$DCHS1, levels=c("WT","MUT"))
circ_data$FSIP2 <- factor(circ_data$FSIP2, levels=c("WT","MUT"))
circ_data$LRP1B <- factor(circ_data$LRP1B, levels=c("WT","MUT"))

mutation_genes <- c("TP53", "CDKN2A", "FAT1", "APOB", "LAMA1", "NOTCH1", "PCDHB13", "DCHS1", "FSIP2", "LRP1B")
summary_table <- data.frame(
  Gene = character(),
  Clearance_MUT_Percent = numeric(),
  No_Clearance_MUT_Percent = numeric(),
  Overall_MUT_Percent = numeric(),
  Fisher_p_value = numeric(),
  stringsAsFactors = FALSE
)
for (gene in mutation_genes) {
  subset_data <- circ_data[!is.na(circ_data[[gene]]) & !is.na(circ_data$ctDNA.Dynamics), ]
  tab <- table(subset_data[[gene]], subset_data$ctDNA.Dynamics)
  total_clearance <- sum(tab[, "Clearance"])
  total_no_clearance <- sum(tab[, "No Clearance"])
  total_all <- sum(tab)
  pct_clearance <- if ("MUT" %in% rownames(tab)) 100 * tab["MUT", "Clearance"] / total_clearance else 0
  pct_no_clearance <- if ("MUT" %in% rownames(tab)) 100 * tab["MUT", "No Clearance"] / total_no_clearance else 0
  pct_overall <- if ("MUT" %in% rownames(tab)) 100 * sum(tab["MUT", ]) / total_all else 0
  fisher_p <- if (all(dim(tab) == c(2,2))) fisher.test(tab)$p.value else NA
  summary_table <- rbind(summary_table, data.frame(
    Gene = gene,
    Clearance_MUT_Percent = round(pct_clearance, 1),
    No_Clearance_MUT_Percent = round(pct_no_clearance, 1),
    Overall_MUT_Percent = round(pct_overall, 1),
    Fisher_p_value = signif(fisher_p, 3)
  ))
}
print(summary_table)
```

